---
title: "Assignment 2 - Exercise 1"
author: "201903536"
date: "02 November 2023"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = FALSE) 


## Setup:
# Install/load packages
library("tidyverse")

```

This is an R-Markdown with the solutions for the first exercise of Assignment 2. It simulates a version control workflow using git and GitHub. 

## Task

Define a function with a single argument data, that takes a dataset and performs some input transformation on it. The function returns the transformed data.

### Installing the data

I am using the palmerpenguins dataset. The description of the dataset can be found [here](https://cran.r-project.org/web/packages/palmerpenguins/readme/README.html). The code needed for installing the palmerpenguins dataset from CRAN can be found in the Code Appendix.

```{r install and load palmerpenguins data}
# Uncomment the following line of code to install the data
#install.packages("palmerpenguins")
library(palmerpenguins)

# The palmerpenguins package contains two datasets.
# For this exercise, we are interested in the penguins dataset.
data(package = 'palmerpenguins')

```

The dataset penguins contains the size measurements for adult foraging penguins near Palmer Station, Antarctica. The dataset holds 344 penguins of three different species (source: [palmerpenguins documentation](https://cran.r-project.org/web/packages/palmerpenguins/readme/README.html)).

```{r summary statistics of the variables}
# Inspect the variables in penguins
summary(penguins)

```

The penguins dataset holds 8 variables holding information on the penguins observed, such as the sex, bill length and species. Note that the dataset does have some missing values.

We inspect the first few rows of the dataset.

```{r inspect the data}
# Show the first few rows of the dataset.
head(penguins)

```

As seen in the first rows of penguins printed out above, the dataset is already tidy (meaning that each row corresponds to one observation, each variable has its own column and each value has its own cell) so we can continue with the task.


### Transforming the data 

Potentially, we are only interested in penguins from the island Biscoe. Therefore, we write a function that filters the dataset for penguins found on Biscoe island. Furthermore, we want to get rid of missing values.

We execute the function on the penguins dataset and save the filtered dataset as "filtered_peng". The new dataset "filtered_peng" should only contain penguins from Biscoe island. By summarising the island variable, we can check whether the function has correctly filtered the dataset.

```{r function that filters penguins for penguins on Biscoe island}

# Write a function filtering for Biscoe island penguins
filter_biscoe <- function(dt) {
  filtered_dt <- dt %>%
    filter(island == "Biscoe") %>%
    na.omit
  return(filtered_dt)}
  

# Execute the function with the penguin data
# and save the filtered dataset as "filtered_peng"
filtered_peng <- filter_biscoe(penguins)

# Check the island variable
summary(filtered_peng$island)

```

As we can see, the filtered dataset only contains 163 observations, which all have been found on Biscoe island. 





























## Appendix: All code in this assignment

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE} 


```
